Left-linear rules in adhesive categories

### 2024-02-13 icalp/fusione.tex
varie cose minori

\begin{abstract}
%  When two sequences of steps executed by a computational device can
%  be considered equivalent?  This is a relevant question if one
%  considers concurrent and distributed systems, since the independence
%  of two steps means that they could be executed in any order, and
%  potentially in parallel.
  When can two sequential steps performed by a computing device be
  considered (causally) independent?  This is a relevant question for
  concurrent and distributed systems, since independence means that
  they could be executed in any order, and potentially in parallel.
  % 
  We investigate the issue in the context of adhesive categories, a
  very general instance of the DPO approach to rewriting.\todo{[A] direi: ``...
    in the context of the Double Pushout (DPO) Approach to rewriting in adhesive categories.''}
    More
  precisely, we focus on left-linear rules, which not only consume,
  preserve and generate entities, as linear rules do, but may
  ``merge'' parts of the state.
  % 
  An apparently minimal, yet technically hard enhancement, 
  % 
  % We refine the established notion of sequential independence
  % between steps, and ...
  since the canonical notion of independence used for linear rewriting
  systems does not work out of the box, but it requires the
  restriction to well-behaved classes of left-linear rules.
  %
  Moreover, a standard characterisation of independence that -- in the
  linear case -- allows one to derive a number of properties,
  essential in the development of a theory of concurrency, no longer
  holds.
  %and this requires a substantial revision of the theory.
  %
  The paper performs an in-depth study of the notion of independence
  for left-linear rules: it introduces a novel characterisation of
  independence, identifies well-behaved classes of left-linear
  rewriting systems and provides some fundamental results including a
  Church-Rosser property and the existence of canonical equivalence
  proofs for concurrent computations. These results properly extends
  the class of formalisms that can be modelled in the adhesive
  framework.
\end{abstract}

@@

Among rewriting-based formalisms, those manipulating graph-like structures
proved useful in many settings for representing the dynamics of distributed systems: the 
graph serves to represent the entities and their relations within states, while 
rewriting steps, modifying the graph, model computation steps that result 
in changes to the system state. The DPO (double-pushout) approach~\cite{EhrigPS73}
is nowadays considered a standard one for these structures, thanks to its locality 
(rule application has a local effect on a state, differently from more recent approaches
like SqPO~\cite{CorradiniHHK06}
or PBPO$^+$~\cite{OverbeekER23}) and to its flexibility: the rules allow to 
specify that,  in a rewriting step, some entities in the current state are removed, 
some others are needed for the rewriting step to occur but remain unaltered and 
some new entities can be generated. Concerning concurrency, the chosen notion 
is shift equivalence, and independence is formally captured by what is called the
Church-Rosser theorem of DPO rewriting~\cite[Section~4]{Ehrig78}.

@@

On the one hand, they capture most categories of
interest for rewriting. In particular, we single out a class of
left-linear rewriting systems on graph-like structures that are
well-switching and we show that rewriting systems on graphs with
equivalences are so. 
As a sanity check, we prove that linear rewriting systems on
$\mathcal{M}$-adhesive categories are well-switching.

@@

There is, however, a further problem. In the literature on linear
rewriting systems, a central result is the characterisation of
switch equivalence of sequences of steps based on what are called processes or
consistent permutations~\cite{Handbook, BaldanCHKS06, tobias}.\todo{Handbook, graph processes, tobias}
In turn this is the key
to derive some simple but fundamental properties for linear rewriting
systems, namely

@@

Unfortunately, for left-linear rewriting systems the characterisation
of switch equivalence based on consistent permutations fails: the
existence of a consistent permutation is only necessary but not
sufficient, hence it cannot be used for proving (1) and (2).
%
This leads to asking whether the properties above hold.
We show that actually much of the theory can be retained: consistency
of switching still holds, while globality of independence holds in a
weaker form, conceptually linked to the fact that fusions introduce a
form of disjunctive causality. Finally, we prove that a canonical form
for switching sequences can also be recovered.

@@

We will use $m\colon X\mto Y$ to denote that an arrow $m\colon X\to Y$ belongs to $\mathcal{M}$.

@@

[Proof in \Cref{propUnique-proof}]

@@

\todo{Se inseriamo le figure in fusione.tex si risparmia dello spazio}
  \begin{figure}
    \input{Figure/rules.tex}

@@

  A derivation $\der{D}$ consisting of three steps $\dder{D}_0$,
  $\dder{D}_1$, and $\dder{D}_2$, each applying the corresponding rule
  $\rho_i$, is depicted in Fig.~\ref{fi:derD}. Note that the numbers in $\rho_1$
  were changed consistently to represent the vertical morphisms.
  Steps $\dder{D}_1$ and $\dder{D}_2$ are clearly sequential
  independent via independence pair $(R_1 \to D_2, L_2 \to D_1)$.
  %that map nodes consistently with their numbering.\todo{P: having names for the graphs would probably help}
  This pair can be used to construct a switch $\der{E}$, shown
  in Fig.~\ref{fi:derE}.

@@

\begin{definition}
  \todo{A: ma questa definizione va tolta, no?}
  Consider derivations $\der{D}_i$ for $i \in \interval[0]{n}$
  related by switches
  \begin{center}
    $\der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots \shift{\nu_n}
    \der{D}_n$
  \end{center}
  It will be referred as a \emph{switching sequence}.  
\end{definition}

@@

\begin{restatable}[Local Church-Rosser Theorem]{theorem}{thmChurch}
  \label{thm:church}
  Let $(i_0, i_1)$ be a strong independence pair
  between $\dder{D}_0$ and $\dder{D}_1$. Then $\dder{D}_0$ and
  $\dder{D}_1$ are switchable.
\end{restatable}
[Proof in \Cref{thmChurch-proof}]

The correspondence between sequential independence and switchability
is fundamental.  A very large class of DPO rewriting systems -- along the
lines of \cite{baldan2011adhesivity} -- can be identified where all
independence pairs are strong for left-linear rewriting systems. The details are in 
\Cref{app:fill}. Here it is important only to observe that such class includes $\cat{Set}$ and it is closed by comma and functor category constructions. 
As such, it includes essentially all categories (e.g., presheaves over set) that are typically considered for modelling purposes.

%\rem{Non introdurrei una classe, come detto sopra}{
\begin{definition}[Strong Enforcing rewriting systems]A left-linear DPO rewriting system is \emph{strong enforcing} if every independence pair between two direct derivations is strong. 
\end{definition}

\begin{example}
  \label{ex:diff2}
  In light of \Cref{thm:church}, \Cref{ex:diff1} provides an example
  of an independence pair which is not strong, and thus of a non-strong enforcing
  adhesive rewriting system. Instead every left-linear system in category $\cat{Graph}$ is
  strong enforcing, by the results of Appendix~\ref{app:fill}, summarized in~\Cref{ex:graph-type-bb} and~\Cref{cor:llr-is-strEnf}.
\end{example}

@@

%\rem{Questo e' quanto gia' asserito sopra, mettere in appendice e riferire}

@@

\begin{restatable}{proposition}{propEqui}
  \label{prop:equi}
  Every linear DPO rewriting system $(\X, \R)$ is strong enforcing.
\end{restatable}
[Proof in \Cref{propEqui-proof}]

\subsection{Well-switchable rewriting systems}\label{subsec:verytame}

Even if we work in strong enforcing rewriting systems where sequential
independence ensures switchability, when dealing with left-linear
rules there is a further -- and possibly more serious -- issue, i.e.,
the fact that there can be more than one independence pairs between
the same derivations. This can hinder the core idea of using
sequential independence as a basis of a theory of concurrency for
rewriting systems, since exchanges performed using different
independence pairs may lead to derivations that are not abstraction
equivalent, and thus one would equate computations that should
definitively be taken apart, as shown in the example below.

@@

  As a consequence, $\der{D}$ and $\der{D}'$ are switch equivalent, but
  this is not acceptable when viewing equivalence classes of
  derivations as concurrent computations: in $\der{D}$ the first two
  steps are not sequential independent, while in $\der{D}'$ they are,
  intuitively because in $\der{D}$ rule $\rho_1$ uses the node
  generated by $\rho_0$, while in $\der{D}'$ rule $\rho_1$ uses the
  node which was in the initial graph. From the technical point of
  view, the property of being switch equivalent is not closed by
  prefix, something which prevents to derive a sensible concurrent
  sematics: if we limit to the first two steps, derivations
  $\dder{D}_0; \dder{D}_1$ and $\dder{D}_0'; \dder{D}_1'$ are not switch
  equivalent while $\der{D}$ and $\der{D}'$ would be
\end{example}

@@

[Proof in \Cref{lemVTame}]

@@

\begin{restatable}{proposition}{lemIde}\label{lem:ide}
	Let $\der{D}$ and $\der{D}'$ be two derivations in a linear DPO-rewriting system $(\X, \R)$ with the same length $l$. If $\id{[0,l-1]}:[0,l-1]\to [0,l-1]$ is a consistent permutation, then $\der{D}$ and $\der{D}'$ are abstraction equivalent.  
\end{restatable}
[Proof in \Cref{lemide-proof}]

\begin{restatable}{proposition}{propcoswitch}\label{prop:coswitch}
Let $\der{D}$ and $\der{E}$ be two switch equivalent derivations, then there exists a consistent permutation between them.
\end{restatable}	
[Proof in \Cref{propcoswitch-proof}]

@@

\begin{restatable}{theorem}{thmcoswitch}
Let $\X$ be an $\mathcal{M}$-adhesive category and 
$\der{D}$ and $\der{D}'$ be two derivations in a linear DPO-rewriting system $(\X, \R)$. 
Then the following are equivalent:
\begin{enumerate}
	\item $\der{D}$ and $\der{D}'$ are switch equivalent;
	\item there exists a consistent permutation $\sigma\colon [0,\lgh(\der{D})-1]\to [0,\lgh(\der{D})-1]$ between them.
\end{enumerate}
\end{restatable}
[Proof in \Cref{thmswitch-proof}]

@@

[Proof~\ref{lemIndepGlobalLeft-proof}]

@@

[Proof~\ref{lemSwitchConfluence-proof}]

@@

[Proof~\ref{thNoNeed-proof}]

@@

[Proof~\ref{coCanonical-proof}]

@@

		We have to construct the two dotted arrows in the diagram below.
		\[\xymatrix@C=15pt{L_0 \ar[d]_{z'_0\circ o_0}&& K_0 \ar[d]_{n_1\circ w}\ar[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@{.>}@/^.35cm/[drrr]_(.4){\alpha_1}|(.29)\hole \ar[dr]|(.28)\hole_{y_1\circ e_0} && L_1 \ar@{.>}@/_.35cm/[dlll]^(.4){\alpha_2} \ar[dl]|(.28)\hole^{y_1\circ b_1}& K_1 \ar[d]^{d_2\circ u}\ar[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{g_2\circ a_1}\\G'_1 && \ar[ll]^{z_1} Q_5 \ar[rr]_{z'_1}&& G'_2  && \ar[ll]^{x_2} Q_3 \ar[rr]_{y_2}&& G_2}\]

@@

	\begin{enumerate}
		\item We have to construct the two dotted arrows in the diagram
		below.
		\[\xymatrix@C=15pt{L_0 \ar[d]_{z'_0\circ o_0}&& K_0
			\ar[d]_{n_1\circ w_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
			\ar@{.>}@/^.35cm/[drrr]_(.4){\beta_0}|(.29)\hole
			\ar[dr]|(.28)\hole_{y_1\circ e_0} && L_1
			\ar@{.>}@/_.35cm/[dlll]^(.4){\beta_1}
			\ar[dl]|(.28)\hole^{y_1\circ b_0}& K_1 \ar[d]^{d_1\circ
				u_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{g_2\circ
				a_1}\\G'_1 && \ar@{>->}[ll]^{z_1} Q_5 \ar[rr]_{z'_1}&& G'_2 &&
			\ar@{>->}[ll]^{x_2} Q_3 \ar[rr]_{y_2}&& G_3}\]

@@

		Our aim is to construct the dotted arrow in the following diagram.
		\[\xymatrix@C=15pt{L_1 \ar[d]_{f_0\circ i_0}&& K_1
			\ar[d]_{q_0\circ v_1}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_1
			\ar@{.>}@/^.35cm/[drrr]_(.4){\epsilon_0}|(.29)\hole
			\ar[dr]|(.28)\hole_{s_1\circ j_1} && L_2
			\ar@{.>}@/_.35cm/[dlll]^(.4){\epsilon_1}
			\ar[dl]|(.28)\hole^{s_1\circ \alpha_1}& K_2
			\ar[d]^{\lambda_0\circ \phi_2}\ar@{>->}[l]_{l_2} \ar[rr]^{r_2} &&
			R_2 \ar[d]^{\xi_1\circ \gamma_1}\\G_0 && \ar@{>->}[ll]^{s_0} Q_0
			\ar[rr]_{t_0}&& H'_1 && \ar@{>->}[ll]^{\xi_0} Q_6
			\ar[rr]_{\xi'_0}&& H'_2}\]

@@

\begin{proof}
	To fix the notation, let us depict the derivation $\der{D}=\{\dder{D}_i\}_{i=0}^2$:
	\[\xymatrix@C=15pt{L_0\ar[d]^{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@/^.35cm/[drrr]_(.4){i_0}|(.29)\hole \ar[dr]|(.28)\hole_{h_0} && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.28)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[r]^{r_1} & R_1 \ar[dr]|(.28)\hole_{h_1} \ar@/^.35cm/[drrr]_(.4){j_0}|(.29)\hole  && L_2 \ar@/_.35cm/[dlll]^(.4){j_1} \ar[dl]|(.28)\hole^{m_2}& K_2 \ar[d]^{k_2}\ar@{>->}[l]_{l_2} \ar[rr]^{r_2} && R_2 \ar[d]_{h_2} \\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2 && \ar@{>->}[ll]^{f_2} D_2 \ar[rr]_{g_2}&& G_3 }\]

@@

	We begin switching $\dder{D}_0$ and $\dder{D}_1$ to get the following diagram.
	\[\xymatrix@C=15pt{L_1\ar[d]^{f_0\circ i_1}&& K_1 \ar[d]_{k'_1}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_1 \ar@/^.35cm/[drrr]_(.4){a_0}|(.29)\hole \ar[dr]|(.28)\hole_{h'_1} && L_0 \ar@/_.35cm/[dlll]^(.4){a_1} \ar[dl]|(.28)\hole^{m'_0}& K_0 \ar[d]^{k'_0}\ar@{>->}[l]_{l_0} \ar[r]^{r_0} & R_0 \ar[dr]|(.28)\hole_{g_1\circ i_0} \ar@/^.35cm/[drrr]_(.4){b_0}|(.29)\hole  && L_2 \ar@/_.35cm/[dlll]^(.4){b_1} \ar[dl]|(.28)\hole^{m_2}& K_2 \ar[d]^{k_2}\ar@{>->}[l]_{l_2} \ar[rr]^{r_2} && R_2 \ar[d]_{h_2} \\G_0 && \ar@{>->}[ll]^{f'_0} D'_0 \ar[rr]_{g'_0}&& G'_1  && \ar@{>->}[ll]^{f'_1} D'_1 \ar[rr]_{g'_1}&& G_2 && \ar@{>->}[ll]^{f_2} D_2 \ar[rr]_{g_2}&& G_3 }\]

>>>>

\begin{abstract}
%  When two sequences of steps executed by a computational device can
%  be considered equivalent?  This is a relevant question if one
%  considers concurrent and distributed systems, since the independence
%  of two steps means that they could be executed in any order, and
%  potentially in parallel.
  When can two sequential steps performed by a computing device be
  considered (causally) independent?  This is a relevant question for
  concurrent and distributed systems, since independence means that
  they could be executed in any order, and potentially in parallel.
  % 
  We investigate the issue in the context of the Double Pushout (DPO)
  Approach to rewriting in adhesive categories, a very general setting
  for rewriting.  More precisely, we focus on left-linear rules, which
  not only consume, preserve and generate entities, as linear rules
  do, but may ``merge'' parts of the state.
  % 
  An apparently minimal, yet technically hard enhancement, 
  % 
  % We refine the established notion of sequential independence
  % between steps, and ...
  since the canonical notion of independence used for linear rewriting
  systems does not work out of the box, but it requires the
  restriction to well-behaved classes of left-linear rules.
  %
  Moreover, a standard characterisation of independence that -- in the
  linear case -- allows one to derive a number of properties,
  essential in the development of a theory of concurrency, no longer
  holds.
  %and this requires a substantial revision of the theory.
  %
  The paper performs an in-depth study of the notion of independence
  for left-linear rules: it introduces a novel characterisation of
  independence, identifies well-behaved classes of left-linear
  rewriting systems and provides some fundamental results including a
  Church-Rosser property and the existence of canonical equivalence
  proofs for concurrent computations. These results properly extends
  the class of formalisms that can be modelled in the adhesive
  framework.
\end{abstract}

@@

Among rewriting-based formalisms, those manipulating graph-like
structures proved useful in many settings for representing the
dynamics of distributed systems: the graph serves to represent the
entities and their relations within states, while rewriting steps,
modifying the graph, model computation steps that result in changes to
the system state. The DPO (double-pushout) approach~\cite{EhrigPS73}
is nowadays considered a standard one for these structures, thanks to
its locality (rule application has a local effect on a state,
differently from more recent approaches like
SqPO~\cite{CorradiniHHK06} or PBPO$^+$~\cite{OverbeekER23}) and to its
flexibility: the rules allow to specify that, in a rewriting step,
some entities in the current state are removed, some others are needed
for the rewriting step to occur but remain unaltered and some new
entities can be generated. Concerning concurrency, the chosen notion
is shift equivalence, and independence is formally captured by what is
called the Church-Rosser theorem of DPO
rewriting~\cite[Section~4]{Ehrig78}\cite[Chapter 3, Sec. 3.4]{CorradiniMREHL97}.

@@

On the one hand, they capture most categories of
interest for rewriting. In particular, we identify a class of
left-linear rewriting systems on graph-like structures that are
well-switching and we show that rewriting systems on graphs with
equivalences are so. 
As a sanity check, we prove that linear rewriting systems on
$\mathcal{M}$-adhesive categories are well-switching.

@@

There is, however, a further problem. In the literature on linear
rewriting systems, a central result is the characterisation of
switch equivalence of sequences of steps based on what are called processes or
consistent permutations~\cite{CMR:GP,Handbook,BaldanCHKS06,heindel2009category}.
In turn this is the key
to derive some simple but fundamental properties for linear rewriting
systems, namely

@@

Unfortunately, for left-linear rewriting systems the characterisation
of switch equivalence based on consistent permutations fails,
% : the existence of a consistent permutation is only necessary but
% not sufficient
hence it cannot be used for proving (1) and (2).
%
This leads to asking whether the properties above hold.
We show that actually much of the theory can be retained: consistency
of switching still holds, while globality of independence holds in a
weaker form, conceptually linked to the fact that fusions introduce a
form of disjunctive causality. Finally, we prove that a canonical form
for switching sequences can also be recovered.

@@

We will write $m\colon X\mto Y$ to denote that an arrow $m\colon X\to Y$ belongs to $\mathcal{M}$.

@@

\full{
[Proof in \Cref{propUnique-proof}]
}

@@

@@

  A derivation $\der{D}$ consisting of three steps $\dder{D}_0$,
  $\dder{D}_1$, and $\dder{D}_2$, each applying the corresponding rule
  $\rho_i$, is depicted in Fig.~\ref{fi:derD}. Note that the numbers in $\rho_1$
  were changed consistently to represent the vertical morphisms.
  Steps $\dder{D}_1$ and $\dder{D}_2$ are clearly sequential
  independent via independence pair $(R_1 \to D_2, L_2 \to D_1)$.
  %that map nodes consistently with their numbering.\todo{P: having names for the graphs would probably help}
  A switch of $\der{D}$ along such independence pair is the derivation
  $\der{E}$ in Fig.~\ref{fi:derE}.

@@

% \begin{definition}
%   Consider derivations $\der{D}_i$ for $i \in \interval[0]{n}$
%   related by switches
%   \begin{center}
%     $\der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots \shift{\nu_n}
%     \der{D}_n$
%   \end{center}
%   It will be referred as a \emph{switching sequence}.  
% \end{definition}

@@

\begin{restatable}[Local Church-Rosser Theorem]{theorem}{thmChurch}
  \label{thm:church}
  Let $(i_0, i_1)$ be a strong independence pair
  between $\dder{D}_0$ and $\dder{D}_1$. Then $\dder{D}_0$ and
  $\dder{D}_1$ are switchable.
\end{restatable}
\full{
  [Proof in \Cref{thmChurch-proof}]
}

The correspondence between sequential independence and switchability
is fundamental. We name the class of rewriting systems where this property holds.

\begin{definition}[Strong Enforcing rewriting systems]
  A left-linear DPO rewriting system is \emph{strong enforcing} if
  every independence pair between two direct derivations is strong.
\end{definition}

Along the lines of \cite{baldan2011adhesivity} we can identify a large
class of adhesive categories such that all left-linear rewriting
systems over such categories are strong enforcing. The details are in
\Cref{app:fill}. Here it is important only to remark that such class
includes $\cat{Set}$ and it is closed by comma and functor category
constructions.  As such, it includes essentially all categories (e.g.,
presheaves over set) that are typically considered for modelling
purposes. Notably, it contains the category $\cat{Graph}$ of directed
graphs.

Still, there are adhesive rewriting systems which are not strong enforcing, as shown by the following example, which is however somehow pathological.

\begin{example}[Non-strong enforcing left-linear rewriting system]
  \label{ex:diff2}
  In light of \Cref{thm:church}, \Cref{ex:diff1} provides an example
  of an independence pair which is not strong, and thus of a
  non-strong enforcing adhesive rewriting system.%  Instead every
  % left-linear system in category $\cat{Graph}$ is strong enforcing, by
  % the results of Appendix~\ref{app:fill}, summarized
  % in~\Cref{ex:graph-type-bb} and~\Cref{cor:llr-is-strEnf}.
\end{example}

@@

@@

\begin{restatable}{proposition}{propEqui}
  \label{prop:equi}
  Every linear DPO rewriting system $(\X, \R)$ is strong enforcing.
\end{restatable}
\full{
[Proof in \Cref{propEqui-proof}]
}

\subsection{Well-switchable rewriting systems}
\label{subsec:verytame}

Even if we work in strong enforcing rewriting systems where sequential
independence ensures switchability, when dealing with left-linear
rules there is a further -- and possibly more serious -- issue, i.e.,
the fact that there can be more than one independence pair between
the same derivations. This can hinder the very idea of using
sequential independence as a basis of a theory of concurrency for
rewriting systems, since exchanges performed using different
independence pairs may lead to derivations that are not abstraction
equivalent, and thus one would equate computations that should
definitively be taken apart, as shown in the example below.

@@

  As a consequence, $\der{D}$ and $\der{D}'$ would be switch
  equivalent, but this is not acceptable when viewing equivalence
  classes of derivations as concurrent computations: in $\der{D}$ the
  first two steps are not sequential independent, while in $\der{D}'$
  they are, intuitively because in $\der{D}$ rule $\rho_1$ uses the
  node generated by $\rho_0$ (adding a self loop to it), while in
  $\der{D}'$ rule $\rho_1$ uses the node which was in the initial
  graph. From the technical point of view, the property of being
  switch equivalent is not closed by prefix, something which prevents
  to derive a sensible concurrent sematics: if we limit to the first
  two steps, we obatain derivations $\dder{D}_0 \cdot \dder{D}_1$ and
  $\dder{D}_0' \cdot \dder{D}_1'$ are not switch equivalent while
  $\der{D} = \dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2$ and
  $\der{D}' = \dder{D}_0'\cdot \dder{D}_1' \cdot \dder{D}_2'$ would
  be.
\end{example}

@@

\full{
[Proof in \Cref{lemVTame}]
}

@@

\begin{restatable}{proposition}{lemIde}\label{lem:ide}
	Let $\der{D}$ and $\der{D}'$ be two derivations in a linear DPO-rewriting system $(\X, \R)$ with the same length $l$. If $\id{[0,l-1]}:[0,l-1]\to [0,l-1]$ is a consistent permutation, then $\der{D}$ and $\der{D}'$ are abstraction equivalent.  
\end{restatable}

\full{
[Proof in \Cref{lemide-proof}]
}

\begin{restatable}{proposition}{propcoswitch}\label{prop:coswitch}
Let $\der{D}$ and $\der{E}$ be two switch equivalent derivations, then there exists a consistent permutation between them.
\end{restatable}

\full{
[Proof in \Cref{propcoswitch-proof}]
}

@@

\begin{restatable}{theorem}{thmcoswitch}
Let $\X$ be an $\mathcal{M}$-adhesive category and 
$\der{D}$ and $\der{D}'$ be two derivations in a linear DPO-rewriting system $(\X, \R)$. 
Then the following are equivalent:
\begin{enumerate}
	\item $\der{D}$ and $\der{D}'$ are switch equivalent;
	\item there exists a consistent permutation $\sigma\colon [0,\lgh(\der{D})-1]\to [0,\lgh(\der{D})-1]$ between them.
\end{enumerate}
\end{restatable}

\full{[Proof in \Cref{thmswitch-proof}]}

@@

\full{[Proof~\ref{lemIndepGlobalLeft-proof}]}

@@

\full{[Proof~\ref{lemSwitchConfluence-proof}]}

@@

\full{[Proof~\ref{thNoNeed-proof}]}

@@

\full{[Proof~\ref{coCanonical-proof}]}

@@

		We have to construct the two dotted arrows in the diagram below.
		\[\xymatrix@C=15pt{L_0 \ar[d]_{z'_0\circ o_0}&& K_0 \ar[d]_{n_1\circ w}\ar[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@{.>}@/^.35cm/[drrr]_(.4){\alpha_1}|(.285)\hole \ar[dr]|(.28)\hole_{y_1\circ e_0} && L_1 \ar@{.>}@/_.35cm/[dlll]^(.4){\alpha_2} \ar[dl]|(.28)\hole^{y_1\circ b_1}& K_1 \ar[d]^{d_2\circ u}\ar[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{g_2\circ a_1}\\G'_1 && \ar[ll]^{z_1} Q_5 \ar[rr]_{z'_1}&& G'_2  && \ar[ll]^{x_2} Q_3 \ar[rr]_{y_2}&& G_2}\]

@@

	\begin{enumerate}
		\item We have to construct the two dotted arrows in the diagram
		below.
		\[\xymatrix@C=15pt{L_0 \ar[d]_{z'_0\circ o_0}&& K_0
			\ar[d]_{n_1\circ w_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
			\ar@{.>}@/^.35cm/[drrr]_(.4){\beta_0}|(.285)\hole
			\ar[dr]|(.28)\hole_{y_1\circ e_0} && L_1
			\ar@{.>}@/_.35cm/[dlll]^(.4){\beta_1}
			\ar[dl]|(.28)\hole^{y_1\circ b_0}& K_1 \ar[d]^{d_1\circ
				u_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{g_2\circ
				a_1}\\G'_1 && \ar@{>->}[ll]^{z_1} Q_5 \ar[rr]_{z'_1}&& G'_2 &&
			\ar@{>->}[ll]^{x_2} Q_3 \ar[rr]_{y_2}&& G_3}\]

@@

		Our aim is to construct the dotted arrow in the following diagram.
		\[\xymatrix@C=15pt{L_1 \ar[d]_{f_0\circ i_0}&& K_1
			\ar[d]_{q_0\circ v_1}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_1
			\ar@{.>}@/^.35cm/[drrr]_(.4){\epsilon_0}|(.285)\hole
			\ar[dr]|(.28)\hole_{s_1\circ j_1} && L_2
			\ar@{.>}@/_.35cm/[dlll]^(.4){\epsilon_1}
			\ar[dl]|(.28)\hole^{s_1\circ \alpha_1}& K_2
			\ar[d]^{\lambda_0\circ \phi_2}\ar@{>->}[l]_{l_2} \ar[rr]^{r_2} &&
			R_2 \ar[d]^{\xi_1\circ \gamma_1}\\G_0 && \ar@{>->}[ll]^{s_0} Q_0
			\ar[rr]_{t_0}&& H'_1 && \ar@{>->}[ll]^{\xi_0} Q_6
			\ar[rr]_{\xi'_0}&& H'_2}\]

@@

\begin{proof}
	To fix the notation, let us depict the derivation $\der{D}=\{\dder{D}_i\}_{i=0}^2$:
	\[\xymatrix@C=13pt{L_0\ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@/^.35cm/[drrr]_(.4){i_0}|(.285)\hole \ar[dr]|(.28)\hole_{h_0} && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.28)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[r]^{r_1} & R_1 \ar[dr]|(.28)\hole_{h_1} \ar@/^.35cm/[drrr]_(.4){j_0}|(.285)\hole  && L_2 \ar@/_.35cm/[dlll]^(.4){j_1} \ar[dl]|(.28)\hole^{m_2}& K_2 \ar[d]^{k_2}\ar@{>->}[l]_{l_2} \ar[rr]^{r_2} && R_2 \ar[d]_{h_2} \\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2 && \ar@{>->}[ll]^{f_2} D_2 \ar[rr]_{g_2}&& G_3 }\]

@@

	We begin switching $\dder{D}_0$ and $\dder{D}_1$ to get the following diagram.
	\[\xymatrix@C=15pt{L_1\ar[d]^{f_0\circ i_1}&& K_1 \ar[d]_{k'_1}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_1 \ar@/^.35cm/[drrr]_(.4){a_0}|(.285)\hole \ar[dr]|(.28)\hole_{h'_1} && L_0 \ar@/_.35cm/[dlll]^(.4){a_1} \ar[dl]|(.28)\hole^{m'_0}& K_0 \ar[d]^{k'_0}\ar@{>->}[l]_{l_0} \ar[r]^{r_0} & R_0 \ar[dr]|(.28)\hole_{g_1\circ i_0} \ar@/^.35cm/[drrr]_(.4){b_0}|(.285)\hole  && L_2 \ar@/_.35cm/[dlll]^(.4){b_1} \ar[dl]|(.28)\hole^{m_2}& K_2 \ar[d]^{k_2}\ar@{>->}[l]_{l_2} \ar[rr]^{r_2} && R_2 \ar[d]_{h_2} \\G_0 && \ar@{>->}[ll]^{f'_0} D'_0 \ar[rr]_{g'_0}&& G'_1  && \ar@{>->}[ll]^{f'_1} D'_1 \ar[rr]_{g'_1}&& G_2 && \ar@{>->}[ll]^{f_2} D_2 \ar[rr]_{g_2}&& G_3 }\]

	By \Cref{def:switch} we also know that $m_0=f'_0\circ a_1$ and $h_1=g'_1\circ a_0$. Now, let us switch $S_{i_0, i_1}(\dder{D}_0)$ and $\dder{D}_2$. In this case we obtain:
	\[\xymatrix@C=15pt{L_1\ar[d]^{f_0\circ i_1}&& K_1 \ar[d]_{k'_1}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_1 \ar@/^.35cm/[drrr]_(.4){c_0}|(.285)\hole \ar[dr]|(.28)\hole_{h'_1} && L_2 \ar@/_.35cm/[dlll]^(.4){c_1} \ar[dl]|(.28)\hole^{f'_1\circ b_1}& K_2 \ar[d]^{k'_2}\ar@{>->}[l]_{l_2} \ar[r]^{r_2} & R_2 \ar[dr]|(.28)\hole_{h'_2} \ar@/^.35cm/[drrr]_(.4){d_0}|(.285)\hole  && L_0 \ar@/_.35cm/[dlll]^(.4){d_1} \ar[dl]|(.28)\hole^{\hat{m}_0}& K_0 \ar[d]^{\hat{k}_0}\ar@{>->}[l]_{l_0} \ar[rr]^{r_0} && R_0 \ar[d]_{g_2\circ b_0} \\G_0 && \ar@{>->}[ll]^{f'_0} D'_0 \ar[rr]_{g'_0}&& G'_1  && \ar@{>->}[ll]^{\hat{f}_1} \hat{D}_1 \ar[rr]_{\hat{g}_1}&& G'_2 && \ar@{>->}[ll]^{f'_2} D'_2 \ar[rr]_{g'_2}&& G_3 }\]
	Using again the definition of switch, we know that $m'_0=\hat{f}_1\circ d_1$ and $h_2=g'_2\circ d_0$. Moreover, by hypothesis $(\X, \R)$ is well switching and the hypothesis of the second point of \Cref{lem:indep-global-left} are satisfied by $\der{D}$. Thus we can characterize $c_0\colon R_1\to \hat{D}_1$ and $c_1\colon L_2\to D'_0$ as the unique arrows fitting in the diagrams below, where the bottom squares are pullbacks.
	\[\xymatrix@R=15pt@C=15pt{&&R_1 \ar@/_1.2cm/[dddl]_(.4){a_0}|(.63)\hole\ar[dl]^{j_0}\ar[dd]^{c'_0} \ar@{.>}[dr]^{c_0}&&&&&L_2 \ar[dl]^{j_1} \ar@/_1.2cm/[dddl]_(.4){z_1}|(.63)\hole\ar[dd]^{c'_1} \ar@{.>}[dr]^{c_1}\\&D_2 \ar@{>->}[dl]^(.4){f_2}&&\hat{D}_1 \ar[dr]^{\hat{g}_1}  &&&D_1 \ar@{>->}[dl]^(.4){f_1}&&D'_0 \ar[dr]^{g'_0}\\G_2 && P\ar[dr]_{p_4} \ar@{>->}[ur]_{p_3}\ar@{>->}[dl]^{p_1} \ar[ul]^{p_2}&&G'_2&G_1 && Q\ar[dr]_{q_4} \ar@{>->}[ur]_{q_3}\ar@{>->}[dl]^{q_1} \ar[ul]^{q_2}&&G'_1\\& D'_1 \ar[ul]^{g'_1}&&D'_2 \ar@{>->}[ur]_{f'_2}&&& D_0 \ar[ul]^{g_0}&&D'_1 \ar@{>->}[ur]_{f'_1}}\]
